The Aquarius Stream Progenitor Was Not a Globular Cluster

### 2012-11-16 2012-aqs/draft-aqs.tex
Model atmospheres section complete

\author{Andrew R. Casey\altaffilmark{1,2}, Stefan C. Keller\altaffilmark{1}, Frebel, Anna\altaffilmark{2}, Gary Da Costa\altaffilmark{1}}
\altaffiltext{1}{Research School of Astronomy \& Astrophysics, Australian National University, Mount Stromlo Observatory, via Cotter Rd, Weston, ACT 2611, Australia}
\altaffiltext{2}{Massachusetts Institute of Technology, Kavli Institute for Astrophysics and Space Research,
77 Massachusetts Avenue, Cambridge, MA 02139, USA}

@@

% Talk about RAVE

@@

We seek to investigate the globular cluster origin claim made by \citet{de_Boer;et-al_2012}. We present a detailed chemical abundance analysis for six confirmed Aquarius stream members observed using the Magellan Inamori Kyocera Echelle spectrograph on the Magellan telescope. Details of the observations and data reduction are outlined in the next section. The data analysis is presented in \S\ref{sec:analysis} and a detailed discussion of these results resides in \S\ref{sec:discussion}, along with an alternative hypothesis for the Aquarius stream origin. In \S\ref{sec:conclusion} we present our conclusions and critical interpretations.

@@

% program observations

% standard star observations

% data reduction using carpy

% normalised each order interactively, stitched them at 5000 Angstroms to provide one full long spectrum

\begin{deluxetable*}{lccccccccc}
\tablecolumns{1}
\tabletypesize{\scriptsize}
\tablecaption{Observations\label{tab:observations}}
\tablehead{
	\colhead{Object} &
	\colhead{$\alpha$} &
	\colhead{$\delta$} &
	\colhead{Observed} &
	\colhead{Airmass} &
	\colhead{Slit Size} &
	\colhead{$t_{exp}$} &
	\colhead{S/N\tablenotemark{a}} &
	\colhead{$V_{rad}$} &
	\colhead{$V_{err}$} \\
 & (J2000) & (J2000) & Date & & (") & (secs) & (px$^{-1}$) & (km s$^{-1}$) & (km s$^{-1}$)
}
\startdata

C2225316-14437	& 22:25:31.7	& $-$14:54:39.6	& 2011-07-30	& 1.033	& 0.7 & \dots & \dots & $-$169.0	& 0.7 \\
C2306265-085103	& 23:06:26.6	& $-$08:51:04.8	& 2011-07-30	& 1.096	& 0.7 & \dots & \dots & $-$239.3	& 0.6 \\
HD41667			& 06:05:03.7	& $-$32:59:36.8	& 2011-03-13	& 1.005	& 1.0 & \dots & \dots & 314.4		& 0.8 \\
HD142948		& 16:00:01.6	& $-$53:51:04.1	& 2011-03-14	& 1.107	& 1.0 & \dots & \dots & 6.8			& 0.4 \\
J221821-183424	& 22:18:21.2	& $-$18:34:28.3	& 2011-07-30	& 1.026	& 0.7 & \dots & \dots & $-$170.5	& 0.5 \\
J223504-152834	& 22:35:04.5	& $-$15:28:34.9	& 2011-07-30	& 1.047	& 0.7 & \dots & \dots & $-$180.9	& 0.7 \\
J223811-104126	& 22:38:11.6	& $-$10:41:29.4	& 2011-07-30	& 1.218	& 0.7 & \dots & \dots & $-$248.4	& 0.7

@@

\subsection{Radial Velocities}
\ref{sec:radial-velocities}
The radial velocity for each star was accurately determined in a two step method. Initially each normalised, stitched spectrum was cross-correlated with a synthetic spectrum of a K0 giant with $T_{eff} = 4500$\,K, $\log{g} = 1.5$, and [M/H] $= -1.0$ across the wavelength range $8450 - 8700$ \AA{}. Velocities found during our cross-correlation are typically within 1\,km s$^{-1}$, and this measurement is used to place each spectrum at its rest wavelength. The rest spectrum was used to measure equivalent widths of approximately N absorption lines in each spectrum (\S\ref{sec:line-measurements}). In addition to an equivalent width and the FWHM of a Gaussian profile which best fits the absorption shape, a measured central wavelength is found for every line. The ratio between the measured central wavelength and the known rest wavelength of the given transition is primarily determined by the stellar radial velocity. Thus, we have N measurements of the stellar radial velocity.

The second step of our radial velocity determination is to find the mean offset and standard deviation from this distribution of line velocity measurements. Figure \ref{fig:line-velocities} shows the line velocities for HD440077 after initially being placed at rest using our cross-correlation velocity of X.XX\,km s$^{-1}$. As expected, the mean offset is small ($-1.1$\,km s$^{-1}$), and our standard deviation is X.XX\,km s$^{-1}$ from N line measurements. This provides us with a final measured radial velocity of $X.XX \pm X.XX$\,km s$^{-1}$. All of our published radial velocities in Table \ref{tab:observations} have been determined in this two-step process.

@@

For the measurement of atomic absorption lines, we employed the line list of \citet{Yong;et-al_2009} with additional transitions of Cr, Sc, Zn, and Sr from \citet{Frebel;et-al_2009}. Molecular line data for CH was taken from \citet{Plez;et-al_2008,Plez;et-al_2009}. We supplemented the list with hyperfine-structure data for Sc and Mn  from the Kurucz compilation \citet{Kurucz;1998}.

We list the atomic data and measured equivalent widths for lines used during this analysis in Table \ref{tab:equivalent-widths}. In order to exclude saturated lines we only used lines with a reduced equivalent widths $\log_{10}{(\mbox{EW}/\lambda)} < -4.5$ in our chemical abundance analysis. A minimum detectable equivalent width was measured as a function of wavelength, and only lines that exceeded a $3\sigma$ detection significance were included. We have verified our equivalent width measurement techniques by measuring equivalent widths for N lines in HD122563 and comparing our measurements with the study of \citet{Norris;et-al_2000}. Excellent agreement is found between the two studies, which is illustrated in Figure \ref{fig:ew-compare}. The mean difference between this study and that of \cite{Norris;et-al_2000} is a negligible $X.X \pm X.X$\,m\AA{}, and no systematic trend is present.

For lines with hyperfine structure, blended transitions or molecular features, we used a spectral synthesis approach. The abundance of a given species was obtained by matching a synthetic spectrum of known abundance to the observed spectrum.

\subsection{Stellar Parameters}

\subsection{Model Atmospheres}

>>>>

\author{Andrew R. Casey\altaffilmark{1,2}, Stefan C. Keller\altaffilmark{1}, Frebel, Anna\altaffilmark{2}, Gary Da Costa\altaffilmark{1}, Alan Alves-Brito\altaffilmark{1}}
\altaffiltext{1}{Research School of Astronomy \& Astrophysics, Australian National University, Mount Stromlo Observatory, via Cotter Rd, Weston, ACT 2611, Australia}
\altaffiltext{2}{Massachusetts Institute of Technology, Kavli Institute for Astrophysics and Space Research,
77 Massachusetts Avenue, Cambridge, MA 02139, USA}

@@

% Talk about RAVE
It is therefore necessary to survey the kinematics of stars in the solar neighbourhood to detect nearby substructures. The RAVE (Radial Velocity Experiment) began such a survey in 2003 and has observed spectra from over 500,000 stars across 17,000 deg$^{2}$. Candidates for RAVE were chosen based only on their apparent magnitude of 9 < I < 13, and therefore are not kinematically biased towards detecting nearby streams. The RAVE data releases have provided radial velocities \citep{Steinmetz;et-al_2006} and estimates of stellar parameters \citep{Zwitter;et-al_2008, Siebert;et-al_2011} now for a subset of these observations.

@@

We seek to investigate the globular cluster origin claim made by \citet{de_Boer;et-al_2012}. We present a detailed chemical abundance analysis for six confirmed Aquarius stream members observed using the Magellan Inamori Kyocera Echelle spectrograph \citep{Bernstein;et-al_2002} on the Magellan telescope. Details of the observations and data reduction are outlined in the following section. The data analysis is presented in \S\ref{sec:analysis} and a detailed discussion of these results resides in \S\ref{sec:discussion}, as well as an alternative hypothesis for the origin of the Aquarius stream. In \S\ref{sec:conclusion} we present our conclusions and critical interpretations.

@@

The most comprehensive sample of Aquarius stream stars is presented in the discovery paper of \citet{Williams;et-al_2011}. Since the globular cluster origin claim of \citet{de_Boer;et-al_2012} is based from only a subset of these stars, we targeted stars largely from the sample of \citet{de_Boer;et-al_2012}. An additional star from the original \citet{Williams;et-al_2011} sample, JXXXXXXXX, was also observed. Program stars were observed in July 2011 and we used standard stars observed for a separate program in March 2011. All standard and program observations were observed using the 1.0" slit, which provides a spectral resolution of $R = 28,000$. The exposure times for our program stars varied per star from X seconds to Y seconds, providing a signal-to-noise in excess of 100 per pixel element at 600 nm for all stars.

The data were reduced using the CarPy pipeline. Standard ThAr arc lamp exposures were taken to provide wavelength calibration. Quartz and milky flat fields at the start of each night were used. No telluric corrections were made as atmospheric absorption does not affect any of the our lines of interest. Each reduced echelle order was interactively normalised using a third order spline. All orders were then stitched together to provide a full spectral coverage from 3800-9400\,\AA{}.

\begin{deluxetable*}{lcccccccccc}
\tablecolumns{1}
\tabletypesize{\scriptsize}
\tablecaption{Observations\label{tab:observations}}
\tablehead{
	\colhead{Designation} &
	\colhead{$\alpha$} &
	\colhead{$\delta$} &
	\colhead{Observed} &
	\colhead{Airmass} &
	\colhead{Seeing} &
	\colhead{$t_{exp}$} &
	\colhead{S/N\tablenotemark{a}} &
	\colhead{$V_{rad}$} &
	\colhead{$V_{hel}$} &
	\colhead{$V_{err}$} \\
 & (J2000) & (J2000) & Date & & (") & (secs) & (px$^{-1}$) & (km s$^{-1}$) & (km s$^{-1}$) & (km s$^{-1}$)
}
\startdata

C2225316-14437	& 22:25:31.7 & $-$14:54:39.6	& 2011-07-30	& 1.033 & \dots & \dots & \dots & $-$169.0	& \dots & 0.7 \\
C2306265-085103	& 23:06:26.6 & $-$08:51:04.8	& 2011-07-30	& 1.096 & \dots & \dots & \dots & $-$239.3	& \dots & 0.6 \\
HD41667			& 06:05:03.7 & $-$32:59:36.8	& 2011-03-13	& 1.005	& \dots & \dots & \dots & 314.4	& \dots & 0.8 \\
HD142948		& 16:00:01.6 & $-$53:51:04.1	& 2011-03-14	& 1.107	& \dots & \dots & \dots & 6.8		& \dots & 0.4 \\
J221821-183424	& 22:18:21.2	& $-$18:34:28.3	& 2011-07-30	& 1.026	& \dots & \dots & \dots & $-$170.5	& \dots & 0.5 \\
J223504-152834	& 22:35:04.5	& $-$15:28:34.9	& 2011-07-30	& 1.047	& \dots & \dots & \dots & $-$180.9	& \dots & 0.7 \\
J223811-104126	& 22:38:11.6	& $-$10:41:29.4	& 2011-07-30	& 1.218	& \dots & \dots & \dots & $-$248.4	& \dots & 0.7

@@

\subsection{Radial Velocities}
\ref{sec:radial-velocities}
The radial velocity for each star was determined in a two step method. Initially each normalised, stitched spectrum was cross-correlated with a synthetic spectrum of a K0 giant with $T_{eff} = 4500$\,K, $\log{g} = 1.5$, and [M/H] $= -1.0$ across the wavelength range $8450 - 8700$ \AA{}. Velocities found during our cross-correlation are typically within 1\,km s$^{-1}$, and this measurement is used to place each spectrum at its approximate rest wavelength. The rest spectrum was used to measure equivalent widths for N absorption lines in each spectrum (\S\ref{sec:line-measurements}). The measurement of equivalent widths involves fitting a Gaussian profile to each absorption feature. Thus, in addition to an equivalent width, the measured central wavelength is found from the Gaussian centroid. The ratio between the measured central wavelength and the laboratory rest wavelength is primarily determined by the stellar radial velocity. Thus, we have N measurements of the stellar radial velocity.

The second step of our radial velocity determination is to find the mean offset and standard deviation from this distribution of line velocity measurements. Figure \ref{fig:line-velocities} shows the line velocities for HD440077 after  being placed at rest using our cross-correlation velocity of X.XX\,km s$^{-1}$. As expected, the mean offset is small ($-1.1$\,km s$^{-1}$), and our standard deviation is X.XX\,km s$^{-1}$ from N line measurements. This provides us with a final measured radial velocity of $X.XX \pm X.XX$\,km s$^{-1}$. This process was repeated for all standards and observations. The radial velocities published in Table \ref{tab:observations} are the final values from this two-step method.

@@

For the measurement of atomic absorption lines, we employed the line list of \citet{Yong;et-al_2009} with additional transitions of Cr, Sc, Zn, and Sr from \citet{Frebel;et-al_2009}. Molecular line data for CH was taken from \citet{Plez;et-al_2008,Plez;et-al_2009}. We further supplemented the list with hyperfine-structure data for Sc and Mn  from the Kurucz compilation \citet{Kurucz;1998}. For lines with hyperfine structure, blended transitions or molecular features, we used a spectral synthesis approach. The abundance of a given species was obtained by matching a synthetic spectrum of known abundance to the observed spectrum.

We list the atomic data and measured equivalent widths for atomic lines used during this analysis in Table \ref{tab:equivalent-widths}. In order to exclude saturated lines we only used lines with a reduced equivalent widths $\log_{10}{(\mbox{EW}/\lambda)} < -4.5$ in our chemical abundance analysis. A minimum detectable equivalent width was measured as a function of wavelength, and only lines that exceeded a $3\sigma$ detection significance were included. We have verified our equivalent width measurement techniques by measuring equivalent widths for N lines in HD122563 and comparing our measurements with the study of \citet{Norris;et-al_2000}. Excellent agreement is found between the two studies, which is illustrated in Figure \ref{fig:ew-compare}. The mean difference between this study and that of \cite{Norris;et-al_2000} is a negligible $X.X \pm X.X$\,m\AA{}, and no systematic trend is present.

\subsection{Model Atmospheres}
We employed the ATLAS9 stellar atmospheres of \citet{Castelli;Kurucz_2003} for this analysis. These models ignore any centre-to-limb spatial variations, assume local thermodynamic equilibrium for any volume within the star as well as overall hydrostatic equilibrium, and assumes no convective overshoot from the stellar photosphere. The stellar parameter spacing for these grid models is 250\,K, 0.5 dex in surface gravity and 0.5 dex in [M/H] and 0.1 dex in [$\alpha$/Fe]. We interpolated the stellar densities, temperatures, electron pressures at all stellar depths between atmospheres using Quickhull algorithm during our analyses. Quickhull is reliant on Delaunay tessellation, which suffers from extremely skewed cells when the grid points vary in size by orders of magnitude -- as $T_{eff}$ values do compared to $\log{g}$ or [M/H]. If unaccounted for, this can result as non-negligible errors in interpolated photospheric quantities across all stellar depths. As such, we scaled each stellar parameter between zero and unity prior to interpolation. During this step, $\log{g}$, [Fe/H] and [$\alpha$/Fe] points were scaled such that we simultaneously interpolate linearly in $T_{eff}$ and logarithmically in all other parameters.

\subsection{Stellar Parameters}
The most recent version of the spectral synthesis code MOOG \citep{Sneden;et-al_1973, Sobeck;et-al_2011} has been used to derive individual line abundances and stellar parameters. This version employs a more accurate treatment of Rayleigh scattering, which is particularly important for transitions blueward of 450\,nm but less relevant for this analysis.
